---
title: "p8105_hw2_jc5929"
output: github_document
date: "2022-10-04"
---

**Question 1**

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
  nyc_transit=read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line,station_name,station_latitude,
         station_longitude,route1:route11,
         entry,vending,entrance_type,ada) %>%
  mutate(entry=recode(entry,'YES'='TRUE','NO'='FALSE'))

```

```{r, eval=FALSE}
nrow(nyc_transit)
ncol(nyc_transit)
```

This dataset contains 19 variables including: line, station_name, 
station_latitude, station_longitude,and the routes number from route1:route11,
entry, vending, entrance_type, 
and also ada. It has 1868 rows, and 19 columns.

I used the function *clean_names* from the package *janitor* to clean the 
variables names in the dataset. This step includes formatting the characters 
from upper case letters into lower case letters, and also added underlines.
Next, I used the *select* function to select the specific columns that I want. 
And then used the *mutate* function to change the entry variable from character 
to logical variables.

```{r}
dis=distinct(nyc_transit,line,station_name,.keep_all=TRUE)
nrow(dis)
dis %>%
  filter(ada=='TRUE')
```

```{r}
dis %>%
  filter(vending=='NO')%>%
  filter(entry=='TRUE')
```

There are 465 dstinct stations. And 84 stations are ada compliant. There are 9 
stations that does not have vending machine, and 5 of them allows entry. 
Therefore, 5/9 station entrances / exits without vending allow entrance.

```{r}
nyc_transit_long =
   pivot_longer(nyc_transit,
                route1:route11,
                names_to='route_number',
                values_to='route_name',
                values_transform = as.character,
                names_transform = as.character)
```

```{r}
nyc_transit_long %>%
  distinct(line,station_name,.keep_all=TRUE) %>%
  filter(route_name=='A')%>%
  filter(ada=='TRUE')
  

```

There are 60 different stations serve A train, and 17 of them are ada compliant.


**Question 2**
```{r q2}
trash_wheel=readxl::read_excel("data/Trash Wheel Collection Data.xlsx",
                       sheet=1,range="A2:N550") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls=round(sports_balls,0),
         sports_balls=as.integer(sports_balls)) %>%
  mutate(trash_wheel_type="mr")
```

```{r}
prof_trash_wheel=readxl::read_excel("data/Trash Wheel Collection Data.xlsx",
                                    sheet=2, range="A2:M97") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls=0) %>%
  mutate(trash_wheel_type="prof") %>%
  select(dumpster:chip_bags,sports_balls,homes_powered,trash_wheel_type) %>%
  mutate(dumpster=as.integer(dumpster),year=as.character(year)) 
  
sum_trash=bind_rows(trash_wheel,prof_trash_wheel)

  
  
```

```{r}
  sum(prof_trash_wheel$weight_tons)
  data=filter(trash_wheel,year==2020)
  sum(data$sports_balls)

```
The sum_trash dataset which combines both the trash_wheel dataset and the
prof_trash_wheel dataset. It has `r nrow(sum_trash)`observations, and 
`r ncol(sum_trash)` columns. Some key variables includes the weight_tons, 
plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, 
chip_bags,sports_balls, which gave a category of the trash. The total weight of
trash produced by professor trash wheel is 190.12 tons, and number of sports 
balls collected by Mr.Trash Wheel in 2020 is 856.

**Question 3**
```{r q3, message=FALSE}
pols_month=read_csv("data/pols-month.csv") %>%
  separate(mon,into=c('year','month','day'),sep='-') %>%
  mutate(month=recode(month,`04`='April',`01`='January',`02`='February',
                      `03`='March',`05`='May',`06`='June',`07`='July',
                      `08`='August',`09`='September',`10`='October',
                      `11`='November',`12`='December'))%>%
  mutate(president=ifelse(prez_dem<1,'gop','dem')) %>%
  mutate(year=as.integer(year)) %>%
  select(-day,-prez_gop,-prez_dem)

snp=read_csv("data/snp.csv") %>% 
  separate(date,into=c('month','day','year'),sep='/') %>%
  mutate(year=as.integer(year)) %>%
  mutate(year=ifelse(year<20,year+2000,year+1900)) %>%
  select(-day) %>%
  select(year,month,everything()) %>%
  mutate(month=recode(month,`4`='April',`1`='January',`2`='February',
                      `3`='March',`5`='May',`6`='June',`7`='July',
                      `8`='August',`9`='September',`10`='October',
                      `11`='November',`12`='December'))


unemploy=read_csv("data/unemployment.csv") %>%
  
  pivot_longer(Jan:Dec,
               names_to='month',
               values_to='rate') %>%
  plyr::rename(c("Year"="year"))  %>%
  mutate(month=recode(month,'Apr'='April','Jan'='January','Feb'='February',
                      'Mar'='March','May'='May','Jun'='June','Jul'='July',
                      'Aug'='August','Sep'='September','Oct'='October',
                      'Nov'='November','Dec'='December')) %>%
  mutate(year=as.integer(year))


ps_data=left_join(pols_month,snp)
sum_data=left_join(ps_data,unemploy)  
  
  
                      
```
The pols_month dataset contains the number of the republican governors, 
republican senators, republican representatives, and also the number of 
democratic governors, democratic governors, democratic representatives on the 
associated date. The range of the timeframe starts from 1947 Janurary to 2015 
June. Moreover, the dataset also contains a column indicating whether the 
president is from dem or gop. 

The snp dataset contains the closing value of the S&P stock on the date of 
observation.

The unemploy dataset indicates the unemployment rate on each month of the 
associated year.

The final dataset sum_data is after joing the above three dataset. It has 
`r nrow(sum_data)`observations, and `r ncol(sum_data)`columns. The range of 
years is from 1947 to 2015. Key variables includes, president, rep_dem, sen_dem,
gov_dem, rep_gop, sen_gop, gov_gop. 

