---
title: "p8105_hw2_jc5929"
output: github_document
date: "2022-10-04"
---


```{r}
library(tidyverse)
library(dplyr)
library(knitr)
  nyc_transit=read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  select(line,station_name,station_latitude,
         station_longitude,route1:route11,
         entry,vending,entrance_type,ada) %>%
  mutate(entry=recode(entry,'YES'='TRUE','NO'='FALSE'))

```

```{r, eval=FALSE}
nrow(nyc_transit)
ncol(nyc_transit)
```

This dataset contains variables: line, station name, station latitude, station
longitude,
and the routes number from route1 to route11, the entry, vending, entrance type, 
and also ADA compliance. It has 1868 rows, and 19 columns.

I used the function *clean_names* from the package *janitor* to clean the 
variables 
names in the dataset. This step includes formatting the characters from upper 
case 
letters into lower case letters, and also added underlines.
Next, I used the *select* function to select the specific columns that I want. 
And then used the *mutate* function to change the entry variable from characrer 
to logical variables.

```{r}
dis=distinct(nyc_transit,line,station_name,.keep_all=TRUE)
nrow(dis)
dis %>%
  filter(ada=='TRUE')
```

```{r}
dis %>%
  filter(vending=='NO')%>%
  filter(entry=='TRUE')
```

There are 465 dstinct stations. And 84 stations are ada compliant. There are 9 
stations that does not have vending machine, and 5 of them allows entry. 
Therefore, 5/9 station entrances / exits without vending allow entrance.

```{r}
nyc_transit_long =
   pivot_longer(nyc_transit,
                route1:route11,
                names_to='route_number',
                values_to='route_name',
                values_transform = as.character,
                names_transform = as.character)
```

```{r}
nyc_transit_long %>%
  distinct(line,station_name,.keep_all=TRUE) %>%
  filter(route_name=='A')%>%
  filter(ada=='TRUE')
  

```

There are 60 different stations serve A train, and 17 of them are ada compliant.

```{r q2}
trash_wheel=readxl::read_excel("data/Trash Wheel Collection Data.xlsx",
                       sheet=1,range="A2:N550") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls=round(sports_balls,0),
         sports_balls=as.integer(sports_balls)) %>%
  mutate(trash_wheel_type="mr")
```
#filter(!is.na(dumpster))
```{r}
prof_trash_wheel=readxl::read_excel("data/Trash Wheel Collection Data.xlsx",
                                    sheet=2, range="A2:M97") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls=0) %>%
  mutate(trash_wheel_type="prof") %>%
  select(dumpster:chip_bags,sports_balls,homes_powered,trash_wheel_type) %>%
  mutate(dumpster=as.integer(dumpster),year=as.character(year)) 
  
sum_trash=bind_rows(trash_wheel,prof_trash_wheel)

  
  
```

```{r}
  sum(prof_trash_wheel$weight_tons)
  data=filter(trash_wheel,year==2020)
  sum(data$sports_balls)

```

```{r q3}
pols_month=read_csv("data/pols-month.csv") %>%
  separate(mon,into=c('year','month','day'),sep='-') %>%
  mutate(month=recode(month,`04`='April',`01`='January',`02`='February',
                      `03`='March',`05`='May',`06`='June',`07`='July',
                      `08`='August',`09`='September',`10`='October',
                      `11`='November',`12`='December'))%>%
  mutate(president=ifelse(prez_dem<1,'gop','dem')) %>%
  mutate(year=as.integer(year)) %>%
  select(-day,-prez_gop,-prez_dem)

snp=read_csv("data/snp.csv") %>% 
  separate(date,into=c('month','day','year'),sep='/') %>%
  mutate(year=as.integer(year)) %>%
  mutate(year=ifelse(year<20,year+2000,year+1900)) %>%
  select(-day) %>%
  select(year,month,everything()) %>%
  mutate(month=recode(month,`4`='April',`1`='January',`2`='February',
                      `3`='March',`5`='May',`6`='June',`7`='July',
                      `8`='August',`9`='September',`10`='October',
                      `11`='November',`12`='December'))


unemploy=read_csv("data/unemployment.csv") %>%
  
  pivot_longer(Jan:Dec,
               names_to='month',
               values_to='amount') %>%
  plyr::rename(c("Year"="year"))  %>%
  mutate(month=recode(month,'Apr'='April','Jan'='January','Feb'='February',
                      'Mar'='March','May'='May','Jun'='June','Jul'='July',
                      'Aug'='August','Sep'='September','Oct'='October',
                      'Nov'='November','Dec'='December')) %>%
  mutate(year=as.integer(year))


ps_data=left_join(pols_month,snp)
sum_data=left_join(ps_data,unemploy)  
  
  
                      
```


